---
title: "Lab1Intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab1Intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab1Intro)
```

# Package Description
[Repository on Github](https://github.com/matt-m-herndon/MATH5793_Lab1Intro)

This package was developed as part of Lab 1 for calculating several fundamental statistical estimators for a given 2d dataframe, passed in as a parameter. The package supports the following:

* The sample mean, given by: 
    $$\bar{x}_k=\frac{1}{N}\sum_{j=1}^{N}x_{jk}$$
* The sample (biased) covariance matrix, leveraging the sample mean and given by: 
    $$S_{ik}=\frac{1}{N}\sum_{j=1}^{N}(x_{ji}-\bar{x}_i)(x_{jk}-\bar{x}_k)$$
* Lastly, the sample correlation matrix, derived from covariance and given by: 
    $$r_{ik}=\frac{S_{ik}}{\sqrt{S_{ii}}\sqrt{S_{kk}}}$$
    
These functions all expect a two dimensional dataframe as an argument.

### Task 3

#### a. JW (1-1,4,5) pages 7,8
$$\bar{x}_k=\frac{1}{N}\sum_{j=1}^{N}x_{jk} \quad \text{where:}\ \ k=1,2,\ldots, p$$
$$S_{ik}=\frac{1}{N}\sum_{j=1}^{N}(x_{ji}-\bar{x}_i)(x_{jk}-\bar{x}_k) \quad \text{where:}\ \ i=1,2,\ldots, p,\ \ k=1,2,\ldots, p$$
$$r_{ik}=\frac{S_{ik}}{\sqrt{S_{ii}}\sqrt{S_{kk}}} \quad \text{where:}\ \ i=1,2,\ldots, p,\ \ k=1,2,\ldots, p$$

#### b. JW (3-27) Page 139
$$\bar{\textbf{x}}_k=\frac{1}{N}{\textbf{X}}'{\textbf{1}} \\ \textbf{S} = \frac{1}{n-1}{\textbf{X}}'\left({\textbf{I}} - \frac{1}{n}{\textbf{11}}'\right){\textbf{X}}$$

#### c. JW (3-28)
$$\underset{(\textbf{p}\times \textbf{p})}{\textbf{D}^{-1/2}} = 
\begin{bmatrix} 
\sqrt{s_{11}} & 0 & \cdots & 0 \\
0 & \sqrt{s_{22}} & \cdots & 0 \\
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \cdots & \sqrt{s_{pp}}
\end{bmatrix} \\$$

#### d. JW (3-29)
$$\textbf{R}=\textbf{D}^{-1/2}\textbf{SD}^{-1/2}$$
